
---
title: "Wine Data Exploration"
author: "Alex"
date: "February 23, 2018"
output: 
  html_document:
  toc: true
  toc_depth: 4
  number_sections: true
  theme: united
  hightlight: tango
  css: my.css
---



### Import Data and Libraries

```{r echo=FALSE, message=FALSE, warning=FALSE}
### Load package

library(ggplot2)
library(dplyr)
library(xlsx)
library(reshape)
library(scales)
library(memisc)
library(GGally)
library(caret)
library(ROCR)


wine = read.csv(file='totalWine.csv')
wine = wine[, -1]

```


> This data consists of 1,599 red and 4,898 white wines sampled in 2009. 
Each sample contains 12 features, plus a quality rating between 0 and 10. 
The basic componenets measured were acidity, sweetness, saltiness, sulphurity, 
pH, alcohol% plus an identification of red or white. 


# Univariate Plots

```{r echo=TRUE}


dim(wine)

```


```{r echo=TRUE}

str(wine)

```



```{r}

summary(wine)

```

> Upon first glance: 
> Featues 1-6, 9-11 appear to be skewed positively. 
> Featues 7 and 12(quality score) appear slightly negatively skewed while 
feature 8 (density) appears uniformly distributed. Lets look a little closer at
each of these features individually. 



```{r, echo=FALSE}

ggplot(data=wine, aes(x=fixed.acidity)) +
  geom_histogram(fill='lightblue', 
                 color='black', bins=100) +
  xlab('Fixed Acidity (g /'~dm^3~')') +
  scale_x_continuous(breaks=seq(3, 16, 1)) +
  #geom_vline(xintercept = mean(wine$fixed.acidity), color='red', size=2) +
  #geom_vline(xintercept = median(wine$fixed.acidity), color='orange', size=2) +
  ylab('Count') +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", 
                                                 scientific=FALSE)) +
  ggtitle('Distribution of Fixed Acidity') +
  theme(plot.title = element_text(hjust = 0.5, size=15))


```

> The distribution is sligthly positively skewed.


```{r echo=FALSE}

summary(wine$fixed.acidity)

```






```{r echo=FALSE}


ggplot(data=wine, aes(x=volatile.acidity)) +
  geom_histogram(fill=I('#FF6666'), 
                 color='black', bins=100) +
  xlab('Volatile Acidity (g /'~dm^3~')') +
  scale_x_continuous(breaks=seq(0, 1.6, 0.2)) +
  #geom_vline(xintercept = mean(wine$fixed.acidity), color='red', size=2) +
  #geom_vline(xintercept = median(wine$fixed.acidity), color='orange', size=2) +
  ylab('Count') +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", 
                                                 scientific=FALSE)) +
  ggtitle('Distribution of Volatile Acidity') +
  theme(plot.title = element_text(hjust = 0.5, size=15))


```

> It looks like there are a couple of outliers


```{r echo=FALSE}
summary(wine$volatile.acidity)
```

```{r echo=TRUE}
quantile(wine$volatile.acidity, probs = 0.999)
```


> The max value lies over twice IQR distance away from the 99.9th percentile
value. 








```{r echo=FALSE}

ggplot(data=wine, aes(x=citric.acid)) +
  geom_histogram(fill=I('#FFFF66'), color='black', bins=100)+ #binwidth = 0.05) +
  xlab('Citric Acid (g /'~dm^3~')') +
  scale_x_continuous(breaks=seq(0, 1.7, 0.2)) +
  #geom_vline(xintercept = mean(wine$fixed.acidity), color='red', size=2) +
  #geom_vline(xintercept = median(wine$fixed.acidity), color='orange', size=2) +
  ylab('Count') +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", 
                                                 scientific=FALSE)) +
  ggtitle('Distribution of Citric Acid') +
  theme(plot.title = element_text(hjust = 0.5, size=15))



```

> The distribution of citric acid appears slightly normally distributed about
its mean. 




```{r echo=FALSE}

ggplot(data=wine, aes(x=residual.sugar)) +
  geom_histogram(fill=I('#CC9966'), color='black', bins=100) +#binwidth = 0.5) +
  xlab('Residual Sugar (g /'~dm^3~')') +
  scale_x_continuous(breaks=seq(0, 66, 8)) +
  #geom_vline(xintercept = mean(wine$fixed.acidity), color='red', size=2) +
  #geom_vline(xintercept = median(wine$fixed.acidity), color='orange', size=2) +
  ylab('Count') +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", 
                                                 scientific=FALSE)) +
  ggtitle('Distribution of Residual Sugar') +
  theme(plot.title = element_text(hjust = 0.5, size=15))





```

> Theres a much wider range of sugar densities than any of the acidic densities.
The max value for residual sugar is approximately 11x greater than the mean. 
The acidic values saw max values approximately 2 - 3 times greater than their 
means. Lets take the log10 distribution of residual sugar.


```{r echo=FALSE, message = FALSE, warning=FALSE}

#,
ggplot(data=wine, aes(x=residual.sugar)) +
  geom_histogram(fill = I('#CC9966'), color='black', bins=100)+ #binwidth = 0.01) +
  xlab('Residual Sugar (g /'~dm^3~')') +
  scale_x_continuous(breaks=c(-2, 0, 2, 4, 8, 24, 66), trans=log10_trans()) +
  #geom_vline(xintercept = mean(wine$fixed.acidity), color='red', size=2) +
  #geom_vline(xintercept = median(wine$fixed.acidity), color='orange', size=2) +
  ylab('Count') +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", 
                                                 scientific=FALSE)) +
  ggtitle('Distribution of Log10Residual Sugar') +
  theme(plot.title = element_text(hjust = 0.5, size=15))



```

> Here we see that there is a slight bimodal distribution with peaks at around 
2 g/dL^3 and at 8 g/dL^3. The outliers above 24 g/dm^3 are more visible after
the Log10 transformation.

```{r echo=FALSE}
summary(wine$residual.sugar)
```





```{r echo=FALSE}


ggplot(data=wine, aes(x=chlorides)) +
  geom_histogram(fill=I('#CCCFFF'), color='black', bins=100)+#binwidth = 0.005) +
  xlab('Chlorides (g /'~dm^3~')') +
  scale_x_continuous(breaks=seq(0, 0.62, 0.05)) +
  #geom_vline(xintercept = mean(wine$fixed.acidity), color='red', size=2) +
  #geom_vline(xintercept = median(wine$fixed.acidity), color='orange', size=2) +
  ylab('Count') +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", 
                                                 scientific=FALSE)) +
  ggtitle('Distribution of Chrlorides') +
  theme(plot.title = element_text(hjust = 0.5, size=15))


```

> This distribution looks very similar to that of the Residual Sugar, except 
with more outliers. 

```{r echo=FALSE}
summary(wine$chlorides)

```




```{r echo=FALSE}


ggplot(data=subset(wine,!is.na(chlorides)), aes(x=chlorides)) +
  geom_histogram(fill=I('#CCCFFF'), color='black', bins=100)+#binwidth = 0.01) +
  xlab('Chlorides (g /'~dm^3~')') +
  scale_x_continuous(trans = log10_trans(), breaks=c(seq(0, 0.12, 0.02), 0.01,
                                                    0.16, 0.24, 0.36, 0.48, 0.60)) +
  #geom_vline(xintercept = mean(wine$fixed.acidity), color='red', size=2) +
  #geom_vline(xintercept = median(wine$fixed.acidity), color='orange', size=2) +
  ylab('Count') +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", 
                                                 scientific=FALSE)) +
  ggtitle('Distribution of Log10 Chrlorides') +
  theme(plot.title = element_text(hjust = 0.5, size=15))


```

> After taking transforming the distribution, the outlier values appear more 
amplified. 



```{r echo=FALSE}

ggplot(data=wine, aes(x=free.sulfur.dioxide)) +
  geom_histogram(fill=I('#33FF99'), color='black', bins=100) + #binwidth = 5) +
  xlab('Free Sulfur Dioxide (mg /'~dm^3~')') +
  scale_x_continuous() +
  #geom_vline(xintercept = mean(wine$fixed.acidity), color='red', size=2) +
  #geom_vline(xintercept = median(wine$fixed.acidity), color='orange', size=2) +
  ylab('Count') +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", 
                                                 scientific=FALSE)) +
  ggtitle('Distribution of Free Sulfur Dioxide') +
  theme(plot.title = element_text(hjust = 0.5, size=15))


```

```{r echo=FALSE}

summary(wine$free.sulfur.dioxide)

```


```{r echo=TRUE}
quantile(wine$free.sulfur.dioxide, probs = 0.999)

```


> Again, the distribution looks like a decay function, with serveral extreme outliers. And the maxium value is over triple IQR distance away from the 99.9th 
percentile value. 




```{r echo=FALSE}

ggplot(data=wine, aes(x=total.sulfur.dioxide)) +
  geom_histogram(fill=I('#33FF99'), color='black', bins=100) + #binwidth = 5) +
  xlab('Total Sulfur Dioxide (mg /'~dm^3~')') +
  scale_x_continuous() +
  #geom_vline(xintercept = mean(wine$fixed.acidity), color='red', size=2) +
  #geom_vline(xintercept = median(wine$fixed.acidity), color='orange', size=2) +
  ylab('Count') +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", 
                                                 scientific=FALSE)) +
  ggtitle('Distribution of Total Sulfur Dioxide') +
  theme(plot.title = element_text(hjust = 0.5, size=15))


```

```{r echo=FALSE}

summary(wine$total.sulfur.dioxide)

```

> In this case, the distribution is slightly bimodal with several outliers. 




```{r echo=FALSE}

ggplot(data=wine, aes(x=density)) +
  geom_histogram(fill=I('#0000FF'), color='white', bins=100) + #binwidth = 5) +
  xlab('Water (g /'~cm^3~')') +
  scale_x_continuous(limits = c(0.96, 1.04),  breaks = seq(0.96, 1.04, 0.01)) +
  #geom_vline(xintercept = mean(wine$fixed.acidity), color='red', size=2) +
  #geom_vline(xintercept = median(wine$fixed.acidity), color='orange', size=2) +
  ylab('Count') +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", 
                                                 scientific=FALSE)) +
  ggtitle('Distribution of Water Density') +
  theme(plot.title = element_text(hjust = 0.5, size=15))


```


```{r}

summary(wine$density)

```


> Water density is cocentrated between 0.99 and 1 g/cm^3



```{r echo=FALSE}

ggplot(data=wine, aes(x=pH)) +
  geom_histogram(fill=I('#0000FF'), color='white', bins=100) + #binwidth = 5) +
  xlab('pH') +
  scale_x_continuous(breaks=seq(2.5, 4.5, 0.10)) +
  #geom_vline(xintercept = mean(wine$fixed.acidity), color='red', size=2) +
  #geom_vline(xintercept = median(wine$fixed.acidity), color='orange', size=2) +
  ylab('Count') +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", 
                                                 scientific=FALSE)) +
  ggtitle('Distribution of pH') +
  theme(plot.title = element_text(hjust = 0.5, size=15))

```

```{r echo=FALSE}

summary(wine$pH)

```

> The distrution of pH is approximately normal. There don't appear to be 
many outliers. 


```{r echo=FALSE}

ggplot(data=wine, aes(x=sulphates)) +
  geom_histogram(fill=I('#0000FF'), color='white', bins=100) + #binwidth = 5) +
  xlab('Sulphates (g /'~dm^3~')') +
  scale_x_continuous() +
  #geom_vline(xintercept = mean(wine$fixed.acidity), color='red', size=2) +
  #geom_vline(xintercept = median(wine$fixed.acidity), color='orange', size=2) +
  ylab('Count') +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", 
                                                 scientific=FALSE)) +
  ggtitle('Distribution of Potassium Sulphate') +
  theme(plot.title = element_text(hjust = 0.5, size=15))

```

```{r echo=FALSE}

summary(wine$sulphates)


```

> This distribution is positively skewed with serveral outliers over double the 
IQR distance from the 75th percentile. 


```{r echo=FALSE}

ggplot(data=wine, aes(x=alcohol)) +
  geom_histogram(fill=I('#FF3300'), color='black', bins=80) + #binwidth = 5) +
  xlab('Alcohol %') +
  scale_x_continuous(breaks = seq(8, 14, 0.5)) +
  #geom_vline(xintercept = mean(wine$fixed.acidity), color='red', size=2) +
  #geom_vline(xintercept = median(wine$fixed.acidity), color='orange', size=2) +
  ylab('Count') +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", 
                                                 scientific=FALSE)) +
  ggtitle('Distribution of Alcohol %') +
  theme(plot.title = element_text(hjust = 0.5, size=15))


```

```{r echo=FALSE}

summary(wine$alcohol)

```

> Alcohol% seems to be distributed in periodic clusters. 


```{r echo=FALSE}

ggplot(data=wine, aes(x=quality)) +
  geom_histogram(fill=I('#FF0099'), color='white', binwidth=1) + #binwidth = 5) +
  xlab('Quality Score') +
  scale_x_continuous(breaks = seq(3,10,1)) +
  #geom_vline(xintercept = mean(wine$fixed.acidity), color='red', size=2) +
  #geom_vline(xintercept = median(wine$fixed.acidity), color='orange', size=2) +
  ylab('Count') +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", 
                                                 scientific=FALSE)) +
  ggtitle('Distribution Quality Score') +
  theme(plot.title = element_text(hjust = 0.5, size=15))


```

```{r echo=TRUE}

table(wine$quality)

```

```{r echo=FALSE}

summary(wine$quality)

```

> The quality score is score is normally distributed with slightly more wines 
tending towards the lower end. 





```{r echo=FALSE}

c = wine %>% 
  dplyr::select(color) %>% 
  dplyr::group_by(color) %>% 
  dplyr::summarise(n = n())


```



```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data=c, aes(x=color)) +
  geom_bar(aes(weight = n, fill=color), name='Wine Color') +
  xlab('Wine Color') +
  ylab('Count') +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", 
                                                 scientific=FALSE)) +   
  ggtitle('Number of Wines for Each Color') +
  theme(plot.title = element_text(hjust = 0.5, size=15))

```




## Univariate Analysis

> This dataset originally consisted of 6,497 rows and 12 columns with more 
than twice as many white wines as reds. 

> There were no missing values. All 12 variables were numerical, including the 
quality score, which was integer valued. I did create a 13th variable, color, 
which describes the color of the wine, and a 14th variable later on which 
asssigns 1 to red wines and 0 to whites. 

> The vast majority of wine samples scored a 5 through 7.
I plotted the distributions of all numerical variables and noticed that 
virtually all of them were positively skewed except for pH, Density and Quality Score, which appeared normally distributed. 

> Alcohol was positively skewed, with a shorter peak count than other variables,
a slow decay and periodic near zero counts approximately 0.75 percentage 
points apart. There were also no perfect or abhorrant tasting wine samples, all quality scores ranged between 3 and 9. All but 35 scored between 4 and 8. 

> The main features I will focus on in this investigation will be: 
Quality Score, Alcohol%, pH, Fixed Acidity, Sulfur Dioxide (Total,Free), 
Density, Chlorides and Residual Sugar. 

> I noticed that virtually all of the postitively skewed histograms contained
extreme outliers, and they were highlighted after taking Log10 transformations.
This begs the question, are these outliers generated by the same wine samples? 

> Lucikly this data set was tidy so it did not require any cleaning. 


# Bivariate Plots 


```{r echo=FALSE}

ggpairs(data=wine[, -c(1, 2, 3, 10, 13)]) +
  ggtitle('Pair Plots Between Wine Features') +
  theme(plot.title = element_text(hjust = 0.5, size=15))


```


> Lets look at pairsiwe correlations more closely. 

```{r echo=TRUE}


cor(wine[, -c(1, 2, 3, 10, 13)])


```

> For the most part, most of this data does not look very hightly correlated (pairwise). 

> A few exceptions include: total sulfur dioxide vs. free sulfur dioxide (0.72)
and water density vs alcohol % (-0.68).



```{r echo=FALSE}

ggplot(data=wine, aes(x=free.sulfur.dioxide, total.sulfur.dioxide)) +
  geom_point(color='#FF66FF', alpha=0.7) +
  geom_smooth(method='lm', se=FALSE) +
  xlab('Free Sulfur Dioxide('~mg/dm^3~')') +
  ylab('Total Sulfur Dioxide ('~mg/dm^3~')') +
  ggtitle('Total Sulfur Dioxide vs. Free Sulfur Dioxide') +
  theme(plot.title = element_text(hjust = 0.5, size=15))



```

 
```{r echo=TRUE}

cor(wine$total.sulfur.dioxide, wine$free.sulfur.dioxide)

```

> It looks like the outilier for both of these variables is attributable to 
one sameple. 



```{r echo=FALSE}


ggplot(data=wine, aes(x=pH, y=fixed.acidity)) +
  geom_point(alpha=0.3, color='blue')+ 
  geom_smooth(method='lm', se=FALSE, col='red') +
  xlab('pH') +
  ylab('Fixed Acidity ('~g/dm^3~')') +
  ggtitle('Fixed Acidity vs. pH') +
  theme(plot.title = element_text(hjust = 0.5, size=15))



```

> This plot looks like it contains two intersecting gropus (perhaps red and 
white wines), which, if independently considered, could make Alcohol% and pH 
more corrleated than they currently appear. The regression line is more heavily
weighed by the lower "chunk". 



```{r echo=TRUE}

cor(wine$pH, wine$fixed.acidity)


```



```{r echo=FALSE}

ggplot(data=wine, aes(x=color, y=alcohol)) +
    geom_jitter(alpha=0.1) +
  geom_boxplot(aes(fill=color), color='blue', alpha=0.6) +
  stat_summary(fun.y='mean', geom='point', color='red', shape=8, size=5) +
  xlab('Wine Color') +
  ylab('Alcohol %') +
  ggtitle('Alcohol% Between Wine Colors') +
  guides(fill=guide_legend('Wine Color')) +
  theme(plot.title = element_text(hjust = 0.5, size=15))

```

> Both wine colors have similar Alcohol% distributions. The IQR of white wines
is sligthly greater than red wines. There are clearly a higher proportion of
low alcohol% wines in the white samples yet both colors were slighlty,
positively skewed.




```{r echo=FALSE}

ggplot(data=wine, aes(x=factor(quality), y=alcohol)) +
  geom_jitter(alpha=0.15) +
  geom_boxplot(color='purple', alpha=0.7) +
  stat_summary(fun.y='mean', geom='point', color='red', fill='red',
               shape=8, size=5) +
  xlab('Quality Score') +
  ylab('Alcohol %') +
  ggtitle('Alcohol% Between Quality Scores') +
  guides(fill=guide_legend('Wine Color')) +
  theme(plot.title = element_text(hjust = 0.5, size=15))

```

```{r echo=TRUE}

cor(wine$alcohol, wine$quality)

```


> Alcohol% is moderately correlated with quality. Top quality samples had
a tendency to be negatively skewed. 




```{r echo=FALSE}

ggplot(data=wine, aes(x=color, y=residual.sugar)) +
  geom_jitter(alpha=0.2) + 
  geom_boxplot(aes(fill=color), color='blue', alpha=0.7) +
  stat_summary(fun.y='mean', geom='point', color='red', shape=8, size=5) +
  xlab('Wine Color') +
  ylab('Residual Sugar ('~g/dm^3~')') +
  scale_y_continuous(breaks=seq(0,70,10)) +
  ggtitle('Boxplots of Residual Sugar Between Wine Colors') +
    guides(fill=guide_legend('Wine Color')) +
  theme(plot.title = element_text(hjust = 0.5, size=15))



```


> Clearly white wines are sweeter on average. The variance of residual sugar is
much smaller for red wines.  



```{r echo=FALSE}

ggplot(data=wine, aes(x=factor(quality), y=residual.sugar)) +
  geom_jitter(alpha=0.2) +
  geom_boxplot(color='purple', alpha=0.7) +
  stat_summary(fun.y='mean', shape=8, size=5, color='red', geom='point') +
  xlab('Quality Score') +
  ylab('Residual Sugar ('~g/dm^3~')') +
  scale_y_continuous(breaks=seq(0,70,10)) +
  ggtitle('Residual Sugars Between Quality Scores') +
    guides(fill=guide_legend('Wine Color')) +
  theme(plot.title = element_text(hjust = 0.5, size=15))



```


```{r echo=TRUE}

cor(wine$residual.sugar, wine$quality)

```

> Residual sugars appear similarly distributed accross quality scores. Sugar and
quality are almost independent to one another. 


```{r echo=FALSE}

ggplot(data=wine, aes(x=color, y=pH)) +
  geom_jitter(alpha=0.2) +
  geom_boxplot(aes(fill=color), color='blue', alpha=0.7) +
  stat_summary(fun.y='mean', shape=8, size=5, color='red', geom='point') +  
  xlab('Wine Color') +
  ylab('pH') +
  scale_y_continuous(breaks=seq(2.5, 4.3, 0.2), limits = c(2.7, 4.1)) +
  ggtitle('Boxplots of pH Between Wine Colors') +
    guides(fill=guide_legend('Wine Color')) +
  theme(plot.title = element_text(hjust = 0.5, size=15))

```

> Clearly red wine has a tendency towards a higher pH, its median pH is greater
than the 75th percentile pH for white wines. 




```{r echo=FALSE}

ggplot(data=wine, aes(x=factor(quality), y=pH)) +
  geom_jitter(alpha=0.2) +
  geom_boxplot(color='purple', alpha=0.7) +
  stat_summary(fun.y='mean', shape=8, size=5, color='red', geom='point') +    
  xlab('Quality Score') +
  ylab('pH') +
  scale_y_continuous(breaks=seq(2.5, 4.3, 0.2), limits = c(2.7, 4.1)) +
  ggtitle('pH Beteween Quality Scores') +
    guides(fill=guide_legend('Wine Color')) +
  theme(plot.title = element_text(hjust = 0.5, size=15))

```

```{r echo=TRUE}

cor(wine$pH, wine$quality)

```


> pH is relatively evenly distributed accross quality scores. 



```{r echo=FALSE}


ggplot(data=wine, aes(x=fixed.acidity, y=density)) +
  geom_point(color= '#FF3366', alpha=0.6)+ 
  geom_smooth(method='lm', se=FALSE) +
  xlab('Fixed Acidity ('~g/dm^3~')') +
  scale_x_continuous(breaks=seq(3,15, 1)) +
  ylab('Density ('~g/cm^3~')') +
  ggtitle('Density vs. Fixed Acidity') +
  theme(plot.title = element_text(hjust = 0.5, size=15))


```



```{r echo=TRUE}

cor(wine$fixed.acidity, wine$density)

```

> Fixed Acidity and Water Density are moderately correlated with one another. 
A couple of the outliers are visible on the Density axis. 



```{r echo=FALSE}


ggplot(data=wine, aes(x=alcohol, y=density)) +
  geom_point(color='#FF9933')+ 
  geom_smooth(method='lm', se=FALSE) +
  xlab('Alcohol%') +
  scale_x_continuous(breaks=seq(7,15, 1)) +
  ylab('Density ('~g/cm^3~')') +
  ggtitle('Density vs. Alcohol%') +
  theme(plot.title = element_text(hjust = 0.5, size=15))



```


```{r echo=TRUE}

cor(wine$density, wine$alcohol)

```

> Density and Alcohol are quite correlated. The discrete, periodic alcohol percentages are visible along the x-axis. This makes sense, naturally more
alcohol would relace water as the ABV increases. 



```{r echo=FALSE}


ggplot(data=wine, aes(x=chlorides, y=residual.sugar)) +
  geom_point(color='orange', alpha=0.2)+ 
  geom_smooth(method='lm', se=FALSE) +
  theme_dark() +
  xlab('Chrlorides('~g/dm^3~')') +
  scale_x_continuous(breaks = seq(0, 0.7, 0.1)) +
  ylab('Density('~g/dm^3~')') +
  ggtitle('Residual Sugar vs. Chrlorides') +
  theme(plot.title = element_text(hjust = 0.5, size=15))




```

> Just like the Fixed Acidity/pH plot, this plot appears to be the intersection
of two seperate "chunks" that are represented by red and white wines. 



## Bivariate Analysis

> The strongest linear relationships are between Water Density/Alcohol% (-0.687 correlation), Density/Fixed Acidity (0.459 correlation) and 
Quality Score/Alcohol% (0.444 correlation), Total Sulfur Dioxide/Free Sulfur 
Dioxide (0.721). 

> There is a significant difference in median alcohol percentages and quality 
scores that is highlighted by the boxplots. 

> White wine samples had a tendency to be much sweeter than red wine samples. 
However it's not very affective when accounting for quality, the median 
sweetness was uniform accross scores. The same can be said about pH.

> I can't say that I was totally surprised by the higher quality samples
containing more alcohol by volume. The negative skews in the higher rated
samples indicate that there are tasty lower ABV wines too. 



# Multivariate Plots



```{r echo=FALSE}

ggplot(data=wine, aes(x=factor(quality), y=alcohol)) +
  geom_boxplot(aes(fill=color)) +
  stat_summary(fun.y='mean',geom='point', color='blue', shape=8, size=5) +
  xlab('Quality Score') +
  ylab('Alcohol %') +
  ggtitle('Boxplots of Alcohol% Between Quality Scores and Wine Colors') +
  guides(fill=guide_legend('Wine Color')) +
  theme(plot.title = element_text(hjust = 0.5, size=15))

```

> Here, we see that accross quality scores, white wines tended to contain 
slightly smaller amounts of alcohol on averge. This is true except for 
score 3 samples which saw white wines containing significantly more alcohol
than reds. 


```{r echo=FALSE}


ggplot(data=wine, aes(x=alcohol, y=density)) +
  geom_point(aes(color=color), alpha=0.4)+
  geom_smooth(method='lm', se=FALSE, color='red') +
  theme_dark() +
  xlab('Alcohol%') +
  scale_x_continuous(breaks=seq(7,15, 1)) +
  coord_cartesian(ylim = c(0.98, 1.02)) +
  ylab('Density ('~g/cm^3~')') +
  facet_wrap(~quality) +
  ggtitle('Density vs. Alcohol%:\nFaceted by Quality Score') +
  theme(plot.title = element_text(hjust = 0.5, size=15))



```

> The linear relationship between Density and Acohol% remains consistent 
accross scores. 


```{r echo=FALSE}

ggplot(data=wine, aes(x=chlorides, y=residual.sugar)) +
  geom_point(aes(color=factor(quality)), alpha=0.4) +
  facet_wrap(~color) +
  geom_smooth(method='lm', se=TRUE, col='red') +
  #  scale_color_brewer(type = 'div',
  #                    palette='Set1', breaks=seq(3,9,1),
  #                   guide = guide_legend(title = 'Quality Score', reverse =F)) +
                                      #override.aes = list(alpha = 0.5, size = 2)))+ 
  theme_dark() +
  scale_color_brewer(palette = 'RdYlGn', name='Alcohol %') +
  xlab('Chlorides (g /'~dm^3~')') +
  scale_x_continuous(breaks = seq(0, 0.6, 0.1)) +
  coord_cartesian(ylim = c(0, 30), xlim=c(0, 0.4)) +
  ylab('Residual Sugar') +
  ggtitle('Residual Sugar vs. Chlorides vs. Alcohol%:\nFaceted by Color') +
  #    guides(fill=guide_legend('Quality Score')) +
  theme(plot.title = element_text(hjust = 0.5, size=15))
  #  scale_fill_manual(name='Quality Score')



```

> In this chart, I wanted to see if there was any relationship between saltiness
and sweetness. I scaled down the coordinates to remove any of the extreme 
outliers. It turns out that the super sweet wines were white while most of 
the salty outliers were from red samples. 



```{r echo=FALSE}


ggplot(data=wine, aes(x=pH, y=fixed.acidity)) +
  geom_point(alpha=0.3, aes(color=factor(quality)))+ 
  geom_smooth(method='lm', se=TRUE, color='red') +
  xlab('pH') +
  ylab('Fixed Acidity ('~g/dm^3~')') +
  facet_wrap(~color) +
  theme_dark() +
  scale_color_brewer(palette = 'RdYlGn', name='Quality Score') +
  ggtitle('Fixed Acidity vs. pH vs Alcohol%:\nFAceted by Color') +
  guides(fill=guide_legend('Wine Color')) +
  theme(plot.title = element_text(hjust = 0.5, size=15))


```

> Wine correlation between pH and Fixed Acidity

```{r echo=FALSE}

cor(wine$pH, wine$fixed.acidity)

```


```{r echo=FALSE}

red = subset(wine, color=='Red')
white = subset(wine, color=='White')
```


> Red Wine correlation between pH and Fixed Acidity

```{r echo=FALSE}

cor(red$pH, red$fixed.acidity)

```

> White Wine correlation between pH and Fixed Acidity

```{r echo=FALSE}

cor(white$pH, white$fixed.acidity)

```

> Factoring for color does improve the correlation betwen pH and Fixed Acidity. 



```{r echo=FALSE}


ggplot(data=wine, aes(x=pH, y=alcohol, color=factor(quality))) +
  geom_point(alpha=0.5, aes(color=factor(quality))) +
  geom_smooth(method='lm', se=FALSE) +
  xlab('pH') +
  ylab('Alcohol %') +
  theme_dark() +
  scale_color_brewer(type='seq', palette = 'Set3',
                     guide=guide_legend(title='Quality')) +
  #scale_x_continuous(breaks = seq(2.5, 4, 0.25)) +
  #geom_smooth(aes(color=color), alpha=0.6, 
  #          stat = 'summary', fun.y='median') +#, name='Color')+     
  #facet_wrap(~color) +
  #scale_color_gradientn(colours=rainbow(20)) +
    ggtitle('Alcohol% vs. pH vs. Quality Score') +
  guides(fill=guide_legend(title='Wine Color')) +
  theme(plot.title = element_text(hjust = 0.5, size=15))



```

```{r echo=TRUE}

cor(wine$alcohol, wine$pH)

```

> We see here that Alcohol% is hardly correlated with pH accross quality
scores. 


```{r echo=FALSE}


ggplot(data=wine, aes(x=free.sulfur.dioxide, total.sulfur.dioxide, 
                      color=color)) +
  geom_point(aes(color=color), alpha=0.7, size=0.5, name='Wine Color') + 
  facet_wrap(~quality) +
  theme_dark() +
  coord_cartesian(xlim = c(0, 150)) +
  xlab('Free Sulfur Dioxide('~mg/dm^3~')') +
  ylab('Total Sulfur Dioxide ('~mg/dm^3~')') +
  ggtitle('Total Sulfur Dioxide vs. Free Sulfur Dioxide') +
  theme(plot.title = element_text(hjust = 0.5, size=15))



```

> Accross the middle 5 quality scores, red wines tended to have a lower Total
Sulfur Dioxide concentrations than white wines. It's difficutlt to make 
an assetion with regards to the other scores, there are too few data points. 




## Multivariate Analysis

> In general, white wines were sweeter, lower in acidity and varied more accross scatterplots. This could simply due to there being almost 4x as many white 
samples as red. I expected to see the quality scores cluster towards certain locations on the scatterplots more consistently (for example in the Alcohol% vs Citric Acid facet grid, higher quality samples did not consistenly appear in 
portions of the grid apart from the lower scores). 

> However in the Density vs. Alcohol% facet grid, there was a noticeable 
tendency for the higher quality samples to be situated towards the lower-right portion of the plot, which is partly explained by higher quality wines having 
higher alcohol%s. 

> After dividing the Fixed Acidity vs. pH plot from the Bivariate Portion into differenct colored plots, it becomes more obvious that the original plot was 
indeed a combination of two intersecting clusters of separate color classes. 
In fact, the correlation between Fixed Acidity and pH elevates when considering 
each wine color seperately. 

> The Total Sulfur Dioxide vs. Free Sulfur Dioxide mainatined their plot shape accross color and quality score afer separation. 

> Below I will perform a quick Logistic Regression Model to try and classify Red wines based on the given data. 




### Extra: Logistic Regression Model (Wine Color Classifier)


```{r echo=FALSE}
wine$red = ifelse(wine$color=='Red', 1, 0)

```



```{r echo=FALSE}

sample = floor(0.7 * nrow(wine))

```


```{r echo=FALSE}

intrain = createDataPartition(y=wine$red, p=0.7, list=FALSE)


```


```{r echo=FALSE}

training = wine[intrain,]

```


```{r echo=FALSE}

testing = wine[-intrain,]

```


```{r echo=FALSE}

m = glm(data=training,red~volatile.acidity+residual.sugar+chlorides+free.sulfur.dioxide+total.sulfur.dioxide+density+alcohol, family=binomial('logit'))
```

> Sulphates, Citric.acid, pH and fixed.acidity were removed due to high p-values
for their parameter estimations. 

```{r echo=FALSE}

summary(m)

```


```{r echo=FALSE}

p = predict(m, newdata = testing, type='response')


```


```{r echo=FALSE}

pr = prediction(p, testing$red)

```



```{r echo=FALSE}

wine.pred = rep('White', length(testing$red))

```


```{r echo=FALSE}

wine.pred = ifelse(p>0.5, 'Red', 'White')
#fitted.results = ifelse(p>0.5, 'Red', 0)

```

> Confusion Matrix

```{r echo=FALSE}

table(testing$color, wine.pred)

```



```{r echo=FALSE}

auc = performance(pr, measure='auc')

```



```{r echo=FALSE}

paste('Accurary: ', auc@y.values[[1]])

```





```{r echo=FALSE}

prf = performance(pr, measure = 'tpr', x.measure = 'fpr')

```


> ROC Curve

```{r echo=FALSE}

plot(prf, title = 'ROC Curve')


```

### Regression Model Analysis

> There was a high degree of accuracy for this sample model, as seen by the ROC 
curve and confusion matrix. I ran 5 70%-30% train-test splits. The model 
accuracy fitted for each of the 5 tests was over 99%. It seems that the 
variables provided in the dataset were sufficient for classifying wine (at 
least for this data set). I wonder if post 2009 data and global wine samples 
would be accurately classified using the model above. 


## Summary Plots

### Plot 1
```{r echo=FALSE}

ggplot(data=wine, aes(x=factor(quality), y=alcohol)) +
  geom_boxplot(aes(fill=color)) +
  stat_summary(fun.y='mean',geom='point', color='blue', shape=8, size=5) +
  xlab('Quality Score') +
  ylab('Alcohol %') +
  ggtitle('Boxplots of Alcohol% Between Quality Scores and Wine Colors') +
  guides(fill=guide_legend('Wine Color')) +
  theme(plot.title = element_text(hjust = 0.5, size=15))

```

> This plot shows that lower quality wines had relatively similar median Alcohol percentages, but after 5, they rose consistently. It seemes that one way to 
increase one's score is to increase the volume of alcohol, no matter the 
intended color of wine. 


### Plot 2
```{r echo=FALSE}

ggplot(data=wine, aes(x=factor(quality), y=residual.sugar)) +
  geom_jitter(alpha=0.2) +
  geom_boxplot(color='purple', alpha=0.7) +
  stat_summary(fun.y='mean', shape=8, size=5, color='red', geom='point') +
  xlab('Quality Score') +
  ylab('Residual Sugar ('~g/dm^3~')') +
  scale_y_continuous(breaks=seq(0,70,10)) +
  ggtitle('Residual Sugars Between Quality Scores') +
    guides(fill=guide_legend('Wine Color')) +
  theme(plot.title = element_text(hjust = 0.5, size=15))



```

> This plot demonstates that overall sweetness was not really a factor in 
the scoring of wine samples. We see that the Residual Sugar distributions are positively skewed accross qualities. In fact, the mean and median sugar amounts
accross scores was almost equivalent. 


### Plot 3

```{r echo=FALSE}


ggplot(data=wine, aes(x=pH, y=fixed.acidity)) +
  geom_point(alpha=0.5, aes(color=factor(quality)))+ 
  geom_smooth(method='lm', se=TRUE, color='red') +
  xlab('pH') +
  ylab('Fixed Acidity ('~g/dm^3~')') +
  facet_wrap(~color) +
  theme_dark() +
  scale_color_brewer(palette = 'RdYlGn', name='Quality Score') +
  ggtitle('Fixed Acidity vs. pH vs Alcohol%:\nFAceted by Color') +
  guides(fill=guide_legend('Wine Color')) +
  theme(plot.title = element_text(hjust = 0.5, size=15))


```

> This plot demonstrates the effectiveness of separting the Fixed Acidity vs. pH scatterplots into separate color classes. The correlation between the two 
variables was increased for both red and white wines after separation. 


## Reflection

> Wine samples with a higher Alcohol % scored better, to my complete and utter surprise (sarcasm). Quality scores remained within a relatively small range(4-8)
with only 35 samples scoring a 3 or 9. In a future analysis, focusing on each 
feature individually would possibly lead to the realization that quality scores 
are arrived at focusinf on different features between wine colors (apart from alcohol%).  

> Luckly I did not suffer through much data cleansing while preparing this 
dataset, it was very, very tidy.

> It was impractical to predict the quality score using a multiple regression 
model, due to its categorical nature. Instead I went for a logistic regression 
model to predict wine color instead of quality score. The model had a very high accuracy after multiple runs of Train-Test splits. Red and white wines really 
do have distinctive characteristics when their compositions are scrutinized. 

> Another difficulty was choosing which plots to pick. The only feature I knew 
about was the acohol content. It can be overwhelming to try and select Bivariate
and Multivariate plots without dowmain knowledge going into the analysis and the
fact that there are > Nc2 possible plot combinations. 

> This dataset could definitley be enriched with Pre and Post 2009 samples and 
winery location. It would be interesting to see how quality 
scores and chemical quantities vary from wine to wine over time, seasonally, 
and accross geographic regions.